---
title: "Brownlows Bets"
author: "Davy d"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "SPRING 2024"    
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
source(file="BrLw.R")
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction- Siulation Study

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

## Our strategy is to build a Function and the use it to answer the three questions. 

The function will be given a class_size by number of students, odds, and Number_of_times brownlow made the bet.  It will give you back how much Brownlows earnings were. We will use it to answer our three questions below. If you can, improve this function by giving default values to some of its inputs. 

### Question 1
In the class and payoff as described, what is the average earnings/losses?

Find the standard deviation and a 95% confidence interval for their earnings
 
  
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlow's winnings (or losings)-->

 
```{r}
nSims=500
saveResults= matrix(rep(0,nSims))
# nSim simulations
for(i in 1:nSims){  
saveResults[i] = BrLw(35,2,637)  #historical did this 637 times
}
hist(saveResults)
print(paste('mean profit= ', round(mean(saveResults))))

sortedResults = sort(saveResults)
#95% confidence interval
lCI = ceiling(sortedResults[0.025*nSims])
uCI = floor(sortedResults[0.975*nSims])
print('95% Confidence Interval, average earnings/losses')

print(c(lCI,uCI))
sdev=sd(saveResults)
print(paste("sd=",round(sdev,2)))
```
 
The mean profit as shown on the graph is around 283 as the mean loss is around 110. (226,340)

### Question 2
What are the fair odds for a class size of 35 students? Manipulation of odds in the function to tune Brownlow's earnings to zero. 


<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->

```{r}
nSims=500
saveResults= matrix(rep(0,nSims))
# nSim simulations
for(i in 1:nSims){  
saveResults[i] = BrLw(35,4.25,637)  #historical did this 637 times
}
hist(saveResults)
print(paste('mean profit= ', round(mean(saveResults))))

sortedResults = sort(saveResults)
#95% confidence interval
lCI = ceiling(sortedResults[0.025*nSims])
uCI = floor(sortedResults[0.975*nSims])
print('95% Confidence Interval, average earnings/losses')

print(c(lCI,uCI))
```
The fair amount that should be paid out is 4.25 (-82, 122).

### Question 3
What size of class makes the 2:1 bet fair?  Use manipulation of number of Students to tune Brownlows earnings to zero.

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlow's betting profits!-->

```{r}
nSims=500
saveResults= matrix(rep(0,nSims))
# nSim simulations
for(i in 1:nSims){  
saveResults[i] = BrLw(28,2,637)  #historical did this 637 times
}
hist(saveResults)
print(paste('mean profit= ', round(mean(saveResults))))

sortedResults = sort(saveResults)
#95% confidence interval
lCI = ceiling(sortedResults[0.025*nSims])
uCI = floor(sortedResults[0.975*nSims])
print('95% Confidence Interval, average earnings/losses')

print(c(lCI,uCI))
```
The class size that gets the odds closest to being fair is 28. ( -47, 100)
# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

The results do make sense, they have enough statistical backing. In order for him to make profit Brownlow has to bet less money on a bigger class size, that he's making a 2:1 ratio for every dollar he is putting in. Zero profit would be a class size of 0. 